Choosing the Choice: Reflections on Modelling Decisions and Behaviour in Demographic Agent-Based Models

### 2017-03-23 article.tex
Reviewers comments

We suggest that inclusion and operationalisation of context are best tackled through an iterative process of development, and evaluation against empirical data. This approach places considerable demands on the modeller in terms of developing the model, but also in exercising their judgement as to what should be included, and identifying where a component should be iterated out. In a sense, this is a similar process to the stepwise approaches commonly used in constructing regression models. Here, however, it is considerably less amenable to automation, both because augmenting a model is time consuming, and because evaluating complex models poses considerable challenges. The challenge of evaluation can be somewhat ameliorated by the use of techniques like sensitivity analysis (see \citet{Thiele2014} for a review of several techniques, and \citet{Oakley2004} for an alternative approach), which can support the modeller in identifying how the individual parameters of a model contribute to overall outputs. This too has limitations, because as we discuss later in relation to heterogeneity in agents (section \ref{heterogeneity-in-decision-strategies}), context can take the form of both process, and parameter.
As a result, the onus remains on the modeller to use their best judgement.

@@

If we limit ourselves further, and consider only how social influence might be modelled, then there are two key questions: what is the unit of influence; and how does influence interact with decision making. It will be useful to address these one at a time, although the two are of course intertwined, since influence is information that affects behaviour. This is also demonstrated by \citet{Mason2007} in their review of models of social influence, where many models treat these aspects in combination.

The flow of influence is synonymous with the flow of information, which has attracted considerable attention in the last few years, with an increasing appreciation of the role of social networks in social influence. Here, rather than the structure over which influence flows, we will focus on what flows over it. 
Where information is transacted, it may be about attitudes, or actions, or outcomes, and can correspond strongly or weakly to the truth. The mechanics of flow can also vary considerably, for example transmission might occur only between direct links, or be passed along many, and may pass in only a single direction.

@@

The other usual interpretation of social information flow is as transmission of information about the state of the world, for example information about expected income, as in models by \citet{Filho2011}, and \citet{Klabunde2014}.  Relating this back to our previous mention of faulty inference, this raises the issue of decision making in an uncertain world, since there is no guarantee that information is accurate. In both models, agents take the average incomes of others in their social network as their expected income were they to relocate. Although some distortion may be introduced by the limited sample, the reported income of other agents is perfectly accurate. 
A potential extension might be to consider inaccuracy arising from cognitive bias; for example, \citet{Mather2000} found that people distorted their recollections in favour of the choices they had made, suggesting that agents might exaggerate their success. Alternatively, \citet{McKenzie2013} found that potential migrants had unrealistically negative expectations about earnings and employment. They suggest this may partly arise from attempting to reduce demands for remittances, i.e. agents underreport their incomes. In either case, this suggests that information transmission is not always simply a passive process.

\citeauthor{McKenzie2013} also suggest that overweighting of negative experiences plays a role, which demonstrates the importance of considering how information is incorporated into the decision process. In this case the suggestion is that another cognitive bias, loss aversion, is a salient feature. This could be reflected by using a decision model explicitly incorporates loss aversion, for example Prospect Theory \citep{Kahneman1979}, or Cumulative Prospect Theory \citep{Tversky1992}. Both variants of Prospect Theory are intended to mimic human decision making, by distorting the perception of high and low probabilities towards certainty, and shifting the value of gains and losses such that large gains are progressively less significant, and losses more so. 
The wider implication is that social information flow may be distorted at the source, or through interpretation, and that agent-based models are uniquely suited to capture this.  Other approaches to decision models are available, for example heuristic methods \citep{Gigerenzer1996}, which emphasise the bounded rationality of human decision making, and Bayesian approaches, which address learning as a component of the decision process \citep{Robbins1964}. In an agent-based modelling context, \citet{Gray2016} have contrasted Bayesian, heuristic, and Cumulative Prospect Theory approaches in a simulation of alcohol misuse disclosure among pregnant women.

>>>>

We suggest that inclusion and operationalisation of context are best tackled through an iterative process of development, and evaluation against empirical data. This approach places considerable demands on the modeller in terms of developing the model, but also in exercising their judgement as to what should be included, and identifying where a component should be iterated out. This is however an expensive exercise, both because augmenting a model is time consuming, and because evaluating complex models poses considerable challenges. The challenge of evaluation can be somewhat ameliorated by the use of techniques like sensitivity analysis (see \citet{Thiele2014} for a review of several techniques, and \citet{Oakley2004} for an alternative approach), which can support the modeller in identifying how the individual parameters of a model contribute to overall outputs. A key benefit to this is that parameters which do not significantly impact the results of a model can be removed. This too has limitations, because as we discuss later in relation to heterogeneity in agents (section \ref{heterogeneity-in-decision-strategies}), context can take the form of both process, and parameter, and sensitivity analysis is directly informative only about the latter.
As a result, the onus remains on the modeller to use their best judgement.

@@

If we limit ourselves further, and consider only how social influence might be modelled, then there are two key questions: what is the unit of influence; and how does influence interact with decision making. It will be useful to address these one at a time, although the two are of course intertwined, since influence is interpreted information that affects behaviour. This is also demonstrated by \citet{Mason2007} in their review of models of social influence, where many models treat these aspects in combination.

The flow of influence and the flow of information are interdependent processes, the latter of which has attracted considerable attention in the last few years, with an increasing appreciation of the role of social networks in social influence. Here, rather than the structure over which influence flows, we will focus on what flows over it. 
Where information is transacted, it may be about attitudes, or actions, or outcomes, and can correspond strongly or weakly to the truth. The mechanics of flow can also vary considerably, for example transmission might occur only between direct links, or be passed along many, and may pass in only a single direction.

@@

The other usual interpretation of social information flow is as transmission of information about the state of the world, for example information about expected income, as in models by \citet{Filho2011}, and \citet{Klabunde2014}.  Relating this back to our previous mention of faulty inference, this raises the issue of decision making in an uncertain world, since there is no guarantee that information is accurate. In both models, agents take the average incomes of others in their social network as their expected income were they to relocate. Although some distortion may be introduced by the limited sample, the reported income of other agents is perfectly accurate. 
A potential extension might be to consider inaccuracy arising from cognitive bias; for example, \citet{Mather2000} found that people distorted their recollections in favour of the choices they had made, suggesting that agents might exaggerate their success. Alternatively, \citet{McKenzie2013} found that potential migrants had unrealistically negative expectations about earnings and employment. They suggest this may partly arise from attempting to reduce demands for remittances, i.e. agents underreport their incomes. In either case, this suggests that information transmission is not always simply a passive process. Influence may also not be exclusively limited to information, and could for example take the form of the provision of help, emotional assistance, or material support - transactions where information is present, but secondary.

\citeauthor{McKenzie2013} also suggest that overweighting of negative experiences plays a role, which demonstrates the importance of considering how information is incorporated into the decision process. In this case the suggestion is that another cognitive bias, loss aversion, is a salient feature. This could be reflected by using a decision model explicitly incorporates loss aversion, for example Prospect Theory \citep{Kahneman1979}, or Cumulative Prospect Theory \citep{Tversky1992}. Both variants of Prospect Theory are intended to mimic human decision making, by distorting the perception of high and low probabilities towards certainty, and shifting the value of gains and losses such that large gains are progressively less significant, and losses more so. 
The wider implication is that social information flow may be distorted at the source, or through interpretation, and that agent-based models are uniquely suited to capture this.  We have focused here on descriptive theories of decision making, precisely because the aim is to model human behaviours as as they are manifested, rather than how they should be. Other approaches to decision models are available, for example heuristic methods \citep{Gigerenzer1996}, which emphasise the bounded rationality of human decision making, and Bayesian approaches, which address learning as a component of the decision process \citep{Robbins1964}. In an agent-based modelling context, \citet{Gray2016} have contrasted Bayesian, heuristic, and Cumulative Prospect Theory approaches in a simulation of alcohol misuse disclosure among pregnant women.
